---
title: "Plot Normalised Survival Fractions"
author: "FiG-T"
date: "`r Sys.Date()`"
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage[T1]{fontenc}
mainfont: SourceSansPro
output: 
  html_document: 
    keep_md: yes
    toc: yes
toc: TRUE
editor_options:
  markdown:
    wrap: 80
  chunk_output_type: inline
---

## Introduction 

Rough plots showing the overall trends in the data for cold survival. The input file for this script is created at the end of 'S01_CalculateNormalisedSurvivalAverages.Rmd'. 

> *Libraries* : Libraries are now listed in the S01_PathsShortcuts.R config file.  This is to make it easier to see which libraries are used across the whole project. 

## Setup

This chunk calls information in from the project-specific setup these files. These contain information on the relative paths to different folders and the colour palettes.

```{r configure-setup, echo=FALSE, message=FALSE, warning=FALSE }
# source all files in configure directory
for (file in list.files(here::here("phenotypes/scripts/S00_ConfigureSetup"))) {
  source(here::here("phenotypes/scripts/S00_ConfigureSetup", file), verbose = FALSE)
}
rm(file)

# set path to data 
path_to_data <- here("phenotypes/data/D01_ColdTolerance/")

# set path to results for this file
path_to_results <- here("phenotypes/results/R01_ColdTolerance/S02_PlotNormalisedSurvivalFractions")
```

## Load Data

```{r load-data}
# load data 
cold_survival_normalised_averages <- readr::read_csv(
  file = here(path_to_data, "normalised_survial_population_averages.csv"), 
  col_names = TRUE, 
  show_col_types = FALSE
)
```

## Exploratory Plots 

### Non-normalised data

Before looking at the normalised data the following plots show changes in the absolute survival number over time. This is to get a sense of how the data is behaving/ whether any biases may be introduced by the normalisation process. 

```{r plot-absolute-survival}
ggplot(
  data = cold_survival_normalised_averages, 
    mapping = aes(
      x = timepoint, 
      y = survical_drop, 
      colour = genotype
    )
  ) +
  geom_point(
    position = position_dodge(0.2), 
    size = 3, 
    show.legend = FALSE
  ) +
  geom_smooth(
    method = "loess", 
    show.legend = FALSE
  ) +
  scale_colour_manual(
    values = aus_fly_palette
  ) +
  labs(
    x = "Timepoint", 
    y = "Absolute Survival (%)"
  ) +
  facet_wrap("genotype") +
  theme_light() +
  muted_strips
```
```{r save-non-normalised-plot}
ggsave(
  filename = here(path_to_results, "F01_NonNormalisedSurvivalFractions.pdf"), 
  plot = last_plot(), 
  bg = "transparent", 
  width = 10, 
  height = 7
)
```

### Checking theorerical maximum normalised increase 

```{r theoretical-increases}
theoretical_max <- cold_survival_normalised_averages |>
  group_by(timepoint) |>
  dplyr::reframe(theory_max = 100/control_ave) |> 
  dplyr::distinct()
```

### Normalised data

```{r plot-normalised-survival}
ggplot(
  data = cold_survival_normalised_averages, 
    mapping = aes(
      x = timepoint, 
      y = survical_drop_normalised
    )
  ) +
  geom_point(
    mapping = aes(
      colour = genotype
    ),
    position = position_dodge(0.2), 
    size = 3, 
    show.legend = FALSE
  ) +
  geom_smooth(
     mapping = aes(
      colour = genotype
    ),
    method = "lm", 
    show.legend = FALSE
  ) +
  scale_colour_manual(
    values = aus_fly_palette
  ) +
  facet_wrap("genotype") +
  geom_point(
    data = theoretical_max |> dplyr::filter(theory_max < 2),
    mapping = aes(
      x = timepoint,
      y = theory_max
    )
  ) +
    labs(
    x = "Timepoint", 
    y = "Normalised Survival"
  ) +
  theme_light() +
  muted_strips
```
> Note: Black dots indicate the 'theoretical maximum' for each timepoint.  I.e.: if 100% survival was observed, what would the normalised value be? 

```{r save-non-normalised-plot}
ggsave(
  filename = here(path_to_results, "F02_NormalisedSurvivalFractionsLinear.pdf"), 
  plot = last_plot(), 
  bg = "transparent", 
  width = 10, 
  height = 7
)
```
