---
title: "Other Exploratory Plots"
author: "FiG-T"
date: "`r Sys.Date()`"
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage[T1]{fontenc}
mainfont: SourceSansPro
output: 
  html_document: 
    keep_md: yes
    toc: yes
toc: TRUE
editor_options:
  markdown:
    wrap: 80
  chunk_output_type: inline
---

## Introduction 

Rough plots checking the impact of 'random' variables on cold survival. The input file for this script is created at the end of 'S01_CalculateNormalisedSurvivalAverages.Rmd'. 

> *Libraries* : Libraries are now listed in the S01_PathsShortcuts.R config file.  This is to make it easier to see which libraries are used across the whole project. 

## Setup

This chunk calls information in from the project-specific setup these files. These contain information on the relative paths to different folders and the colour palettes.

```{r configure-setup, echo=FALSE, message=FALSE, warning=FALSE }
# source all files in configure directory
for (file in list.files(here::here("phenotypes/scripts/S00_ConfigureSetup"))) {
  source(here::here("phenotypes/scripts/S00_ConfigureSetup", file), verbose = FALSE)
}
rm(file)

# set path to data 
path_to_data <- here("phenotypes/data/D01_ColdTolerance/")

# set path to results for this file
path_to_results <- here("phenotypes/results/R01_ColdTolerance/S03_OtherExploratoryPlots")
```

## Load Data

Using the ungrouped survival data: 

```{r load-data}
# load data 
cold_survival_normalised <- readr::read_csv(
  file = here(path_to_data, "normalised_survival.csv"), 
  col_names = TRUE, 
  show_col_types = FALSE
)
```

```{r format-factor-levels}
# set the order of genotypes
cold_survival_normalised$genotype <- factor(
  cold_survival_normalised$genotype, 
  levels = genotype_factor_levels
)

# set the order of populations
cold_survival_normalised$pop <- factor(
  cold_survival_normalised$pop, 
  levels = population_factor_levels
)
```

## Exploratory Plots 

```{r replicate-impact}
# looking across vials 
ggplot(
    data = cold_survival_normalised, 
    mapping = aes(
      x = replicate, 
      y = survival_drop_normalised, 
      colour = genotype
    )
  ) +
  geom_point(
    show.legend = FALSE
  ) +
  scale_colour_manual(
    values = aus_fly_palette
  ) +
  geom_smooth(
    method = "lm", 
    colour = "black", 
    linetype = "dotdash"
  ) +
  facet_grid(
    rows = vars(genotype), 
    cols = vars(timepoint)
  ) +
  labs(
    x = "Replicate (vial number)", 
    y = "Normalised Survivial"
  ) +
  theme_light() +
  muted_strips

```
```{r save-non-normalised-plot}
ggsave(
  filename = here(path_to_results, "F01_NormalisedSurvivalFractionsVsReplicateNumber.pdf"), 
  plot = last_plot(), 
  bg = "transparent", 
  width = 10, 
  height = 7
)
```
This plot suggests that replicate number has a random effect (as expected) on normalised survivial. 

### Plot Initial Temperature Effects 

```{r replicate-impact}
# looking across vials 
ggplot(
    data = cold_survival_normalised , 
    mapping = aes(
      x = initial_temp, 
      y = survival_drop_normalised, 
      colour = genotype
    )
  ) +
  geom_point(
    show.legend = FALSE
  ) +
  scale_colour_manual(
    values = aus_fly_palette
  ) +
  geom_smooth(
    method = "lm", 
    colour = "black", 
    linetype = "dotdash"
  ) +
  labs(
    x = "Starting Temperature (ºC)",
    y = "Normalised Survivial"
  ) +
  theme_light() +
  muted_strips

```
```{r save-non-normalised-plot}
ggsave(
  filename = here(path_to_results, "F02_NormalisedSurvivalFractionsVsStartingTemperature.pdf"), 
  plot = last_plot(), 
  bg = "transparent", 
  width = 10, 
  height = 7
)
```

Looking across all timepoints and genotypes, there also appears to be no strong impacts from starting temperature.

### Plot Tray Effects

```{r replicate-impact}
# looking across vials 
ggplot(
    data = cold_survival_normalised , 
    mapping = aes(
      x = tray, 
      y = survival_drop_normalised, 
      colour = genotype
    )
  ) +
  geom_point(
    show.legend = FALSE, 
    position = position_dodge(0.6)
  ) +
  scale_colour_manual(
    values = aus_fly_palette
  ) +
  geom_smooth(
    method = "lm", 
    colour = "black", 
    linetype = "dotdash"
  ) +
  labs(
    x = "Tray",
    y = "Normalised Survivial"
  ) +
  facet_grid(
    cols = vars(timepoint)
  ) +
  theme_light() +
  muted_strips

```
```{r save-non-normalised-plot}
ggsave(
  filename = here(path_to_results, "F03_NormalisedSurvivalFractionsVsTray.pdf"), 
  plot = last_plot(), 
  bg = "transparent", 
  width = 10, 
  height = 7
)
```

This also confirms that tray does not have an effect on the survival rate. 
